---
title: "Ethnicity , Cost-Distance, and how Civil Wars Spread"
author: "Daniel K Baissa"
date: "3/14/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rgdal)
library(raster)
library(tidyverse)
library(sf)
```

## Intro

## Managing the Shapefiles


I will start by merging the Iraq and Syria shapefiles. I will start by binding the Iraq and Syria shapefiles.

```{r echo=T, message=FALSE, warning=FALSE, results="hide"}

Iraq <-st_read("Shapefiles/Iraq/pc284dt8747.shp")
Syria <-st_read("Shapefiles/Syria/gk468fm4741.shp")

Iraq_and_Syria <- st_union(Iraq, Syria)

ggplot(data = Iraq_and_Syria) +
    geom_sf()
```

```{r echo = T, results = 'hide'}
world_ethnic_groups <-st_read("Shapefiles/GeoEPR-2021/GeoEPR-2021.shp")

world_ethnic_groups <- st_transform(world_ethnic_groups, crs = 4326)

ggplot(data = world_ethnic_groups) +
    geom_sf()
```

```{r}
Iraq_Syria_groups <- world_ethnic_groups %>%
  filter(statename == "Syria" | statename == "Iraq")

ggplot(data = Iraq_and_Syria) +
    geom_sf()+
    geom_sf(data= Iraq_Syria_groups, aes(fill = group) , alpha = .8)
```

```{r}
# st_write(Iraq_Syria_groups, 'Shapefiles/Iraq_Syria_merged/Iraq_Syria_groups.shp')
```

### Rasterizing the Ethnic Data

I theorize that ethnic groups have an effect on a groups ability to hide and maneuver. Groups that are hostile to the rebel group are less likely to hide or aid the insurgents, where as co-ethnics are more likely to hide them (cites).

This means that I will convert the raster data into a cost surface which can be used with our geographic and climitalogical data.

Now Let's create a function that will apply values depending on group relations.

```{r}

shp <- rgdal::readOGR("Shapefiles/Iraq_Syria_merged/Iraq_Syria_groups.shp")


group_cost <- function(in_group = "Sunni Arabs", 
                       out_group = c("Shi'a Arabs", 
                                     "Kurds", 
                                     "Assyrians",
                                     "Turkmen",
                                     "Christians",
                                     "Alawi",
                                     "Druze"),
                       # neutral = NA,
                       shapefile = shp,
                       in_group_bonus = .5,
                       out_group_cost = 5)
  {
  shp@data[["group_value"]] <- 1
  shp@data[["group_value"]][which(shp@data[["group"]] %in% in_group)] <- in_group_bonus
  shp@data[["group_value"]][which(shp@data[["group"]] %in% out_group)] <- out_group_cost
  
  # Rasterizing
  template <- raster(ext = extent(shp), crs = projection(shp))

  rst <- rasterize(shp, template, field = "group_value")
  
  return(rst)
}
```

Now let's create some cost rasters using this function. I will start with the Sunni Arabs since ISIS was a Sunni group

```{r}
sunni <- group_cost(shapefile = shp, in_group_bonus = .25)

plot(sunni)
```

Now I will plot the coalition against ISIS. For now I will simplify them into one cost group, but in the future they will be broken down by each appropreat group.

```{r}
coalition <- group_cost(out_group = "Sunni Arabs", 
                       in_group = c("Shi'a Arabs", 
                                     "Kurds", 
                                     "Assyrians",
                                     "Turkmen",
                                     "Christians",
                                     "Alawi",
                                     "Druze"),
                       shapefile = shp, 
                       in_group_bonus = .75,
                       out_group_cost = 2)

plot(coalition)
```

## Managing the Raster data

Loading in the rasters

```{r}
roads <- raster("rasters/road_rasters/iraq_and_syria_road.tif")
cost_r <- raster("rasters/cropped_masked/cost_without_roads.tif")
```

```{r}

# cropped <- crop(roads, cost_r)
cropped <- crop(cost_r, roads)
cropped <- resample(cropped, roads)
plot(cropped)
plot(roads)

```



Roads are a more efficient way of moving and therefore have a lower cost. I will multiply the roads by 0.05 and then multiply that roads raster to the cost raster.

```{r}
cost_of_roads <- 0.05

roads[roads == 1] <- cost_of_roads

roads[roads == 0] <- 1

plot(roads)

```


```{r}

cost_raster <- cropped*roads

plot(cost_raster)
```

## Calculating ISIS's extent
The first mover in my function is the government. They have resources deployed, ready to counter threats. In this example, I will say the coalition is the government. Again, this is just for now while I create all of these functions and ensure that they are working.



The centers of power for the coalition. Baghdad, Erbil, and Damascus

```{r}
x_coords <- c(44.3661, 43.9930, 36.2765) 
y_coords <- c(33.3152, 36.1901, 33.5138)


Baghdad_Erbil_and_Damascus <- SpatialPoints(coords = cbind(x_coords,y_coords))
```


```{r}
cropped_coalition <- crop(coalition, roads)
cropped_coalition <- resample(cropped_coalition, roads)

cost_raster <- (cropped*roads)*(cropped_coalition)

plot(cost_raster)
```

```{r}
tr <- transition(cost_raster, function(x) 1/mean(x), 16) 
tr <- geoCorrection(tr)

coalition_a <- accCost(tr, Baghdad_Erbil_and_Damascus)

plot(coalition_a, col=colorRampPalette(c("black", "white", "green", "yellow", "orange", "red", "purple"))(255))
plot(log(coalition_a), col=colorRampPalette(c("black", "white", "green", "yellow", "orange", "red", "purple"))(255))

```

Turning the projection cost of the coaltion into a cost raster by inverting it.
```{r}
projection_cost <- (log(coalition_a)*-1) + 20

plot(projection_cost)

```


Now that we have the Coalition's ability to project power, let's look at ISIS, the second mover.

```{r}

cropped_Sunni <- crop(sunni, roads)
cropped_sunni <- resample(cropped_Sunni, roads)

cost_raster <- (cropped*roads)*(cropped_sunni)*projection_cost

plot(cost_raster)

```

### ISIS's cost path

Center's of power for ISIS:

Raqqa, Mayadin, Qa'im

```{r}
x_coords <- c(38.9981, 40.4465, 41.1603, 42.3742, 40.7311)
y_coords <- c(35.9594, 35.0161, 34.3169, 34.1252, 36.0558)


ISIS_centers_of_power <- SpatialPoints(coords = cbind(x_coords,y_coords))
```


```{r}
tr <- transition(cost_raster, function(x) 1/mean(x), 16) 
tr <- geoCorrection(tr)

a <- accCost(tr, ISIS_centers_of_power)

plot(a, col=colorRampPalette(c("black", "white", "green", "yellow", "orange", "red", "purple"))(255))
plot(log(a), col=colorRampPalette(c("blue","green", "yellow",  "red"))(255))

# la <- log(a)

```



