---
title: "Mapping Roads in Iraq and Syria"
author: "Daniel K Baissa"
date: "3/3/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intro

For my dissertation I will use road data in Iraq and Syria. In this document, I will extract and merge all of the requisit shapefiles.

The libraries I will use are:

```{r}
library(rgdal)
library(raster)
```


I will start by merging the Iraq and Syria shapefiles. I will start by binding the Iraq and Syria shapefiles.

```{r}
Iraq <- readOGR("Shapefiles/Iraq", "pc284dt8747")
Syria <- readOGR("Shapefiles/Syria", "gk468fm4741")
Iraq_and_Syria <- bind(Iraq, Syria)
plot(Iraq_and_Syria)
```

Now that the shapefiles have been bound, I can aggregate them.

```{r}
Iraq_Syria_joined <- aggregate(Iraq_and_Syria)
plot(Iraq_Syria_joined)
```

Now that we have the Iraq and Syria area mapped out, I will start adding the roads.

I will read in the in the shapefiles.

```{r}
Iraq_roads <- readOGR("Shapefiles/Iraq/primary_routes", "primary_routes")
Syria_roads <- readOGR("Shapefiles/Syria/roads", "hotosm_syr_roads_polygons")
```

Now I will bind the shapefiles

```{r}
# Iraq_and_Syria_roads <- bind(Iraq_roads, Syria_roads)

plot(Iraq_roads)
plot(Syria_roads)

```

